---
title: "REUDataExploration2"
author: "Courtney Miller"
date: "6/6/2018"
output: html_document
---

```{r setup, include=FALSE}
library(Lock5Data)
library(Lock5withR)
library(ggplot2)
library(readr)
library(dplyr)
data <- read_csv("~/Desktop/CMU REU '18/REUDataExploration/UnCorrectedDataHadoop.csv")
knitr::opts_chunk$set(echo = TRUE)
```

### Breaking up the Data
- The data is initially imported as one large data set, to reduce the number of SQL queeries necessary 
-First I will break the data set into smaller tables based on the author id

```{r, echo=FALSE}
#Note: had to change order of offsets so they matched up with the correct author_id, watch out for this!
authorOffset <- c(-467, -226, -735, 0, 900) 
#authorIds <- c(535, 348705, 790256, 840830, 1442272) - not necessary since we used indexes in the for loop

#Splits large data frame up into smaller ones based on author_id
splittingByAuthor <- split(data, data$author_id)

for(i in seq_along(authorOffset)){
  #Adjusts HoW based of list of tz offsets (be careful about order of list!! make sure it matches order of data frames!)
  splittingByAuthor[[i]] <- mutate(splittingByAuthor[[i]], localHoW = (HoW + as.integer(authorOffset[i]/100)) %% 168)
  
  #Creates HoD and o.h Variables for each author_id
  splittingByAuthor[[i]] <- splittingByAuthor[[i]] %>% mutate(localO.h = ifelse(localHoW %in% 129:137, "1", ifelse(localHoW %in% 105:113, "1", ifelse(localHoW %in% 81:89, "1", ifelse(localHoW %in% 57:65, "1", ifelse(localHoW %in% 33:41, "1", "0")))))) %>% mutate(localHoD = ifelse(localHoW %in% 0:23, (localHoW %% 24) + 24, ifelse(localHoW %in% 144:167, local(HoW %% 24) + 24, localHoW %% 24))) 
}

```

#Function to Create Heat Map and Line Graph

**Heat Map Description:**

- The y-axis in these visualizations represent the hour of day (HoD) when the commit was made. 0-23 are the aggregate hours of the work week, and 24-47 are the aggregate hours for the weekend (i.e. the cell at (2013, 12) represents the total number of commits made in 2013 at 12:00 on a work day).  The x-axis represents the year the commit was made in.  This design choice was suggested by Chris (and I believe originated from Christian) because we want to examine their overall work pattern (do they work during office hours or non-office hours?), and this allows us to see the difference while keeping the scale manageable (the alternative would of been to have all the hours of the week 0-168 and this seemed much less digestible as a visualization).

**Line Graph Description:**

- For these visualizations, the focus is again on the top 5 contributors, except now the visualization will be looking at the number of commits during office hours (9-5 on weekdays) vs non-office hours (non-9-5/weekends).  I chose to visualize this plot with a line graph, where the x-axis is the year and the y-axis is the number of commits.  Both office and non-office hours are plotted on the same graph but in different colors, to help streamline the visualization.

- First I created a new office hour variable (o.h), which will be used for grouping in line graph, 1 = office hours, 0 = non-office hours.  I considered office hours to be from 09:00 - 17:00.  However it should be noted potential time zone issues have not yet been addressed.  

- To see the graphs for an author, enter their identifier into the funtion 'create_both_graphs(authorInfo)'
- ex. '>*create_both_graphs(splittingByAuthor[[1]])*'

```{r, echo=FALSE}
#Funtion takes in an author and creates line line graph plotting office hours vs. non-office hours and longitudinal Heat Map of Working Hours for the author 
create_both_graphs <- function(authori) {
     # Aggregates frequencies and groups by o.h and year
    ai <- authori %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)
    
    # Changes o.h from continuous to discrete column 
    ai$localO.h <- as.factor(ai$localO.h)
    
    # Creates line graph plotting office hours vs. non-office hours
    plot(ggplot(ai, aes(x = year, y = freq)) + geom_line(aes(colour= localO.h)) + labs(title = "Office Hour vs. Non-Office Hour Commits\n", x = "Year", y = "Number of Commits", color = "Time of Commit") +
      scale_color_manual(labels = c("During Non-Office Hours", "During Office Hours"), values = c("blue", "red")))
    
    # Aggregates frequencies and groups by HoD
    aib <- authori %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, localHoD, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)
    
    # Creates longitudinal Heat Map of Working Hours:
    plot(ggplot(aib, aes(year, localHoD)) + geom_tile(aes(fill = freq), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue") + scale_y_continuous(breaks=seq(0, 168, 2)))
}
```