---
title: "REUDataExploration2"
author: "Courtney Miller"
date: "6/6/2018"
output: html_document
---

```{r setup, include=FALSE}
library(Lock5Data)
library(Lock5withR)
library(ggplot2)
library(readr)
library(dplyr)
data <- read_csv("~/Desktop/CMU REU '18/REUDataExploration/UnCorrectedDataHadoop.csv")
knitr::opts_chunk$set(echo = TRUE)
```

### Breaking up the Data
- The data is initially imported as one large data set, to reduce the number of SQL queeries necessary 
-First I will break the data set into smaller tables based on the author id

```{r, echo=FALSE}
authors <- c(`1442272`, `840830`, `535`,`348705`,`790256`)

splittingByAuthor <- split(data, data$author_id)
author1 <- splittingByAuthor$`1442272`
author2 <- splittingByAuthor$`840830`
author3 <- splittingByAuthor$`535`
author4 <- splittingByAuthor$`348705`
author5 <- splittingByAuthor$`790256`
View(author5)
```

### Adjusting for Time Zone
- Now I will take the hour of work (HoW) for each person's commits and adjsut then using the mod equation based (provided by Chris) on their time zone
- The mod equation is:  **localHoW** = (**HoW** + as.integer(*offset*/100) %% 168)
- Note, even if the author has an offset of 0, we still create a localHoW variable for the sake of consistency in later code thats just equal to HoW
```{r, echo= FALSE}

for (authorid in authors) {
  mutate(splittingByAuthor[[authorid]],localHoW = (HoW + as.integer(   **** put in the author's offset here *** /100)) %% 168)
}
#Author 1, offset = 900
author1 <- mutate(author1, localHoW = (HoW + as.integer(900/100)) %% 168)
#Author 2, offset = 0 (so localHoW = HoW)
author2 <- mutate(author2, localHoW = HoW)
#Author 3, offset = -467
author3 <- mutate(author3, localHoW = (HoW + as.integer(-467/100)) %% 168)
#Author 4, offset = -226
author4 <- mutate(author4, localHoW = (HoW + as.integer(-226/100)) %% 168)
#Author 5, offset = -735
author5 <- mutate(author5, localHoW = (HoW + as.integer(-735/100)) %% 168)
```


### Creating HoD and o.h Variables
- For the sake of exploration, we will create an HoD variable for the local time zone.
- We will also create the frequency count variable which we will use to aggregate frequiencies in later steps.

```{r, echo= FALSE}
#Author 1 (local):
author1 <- author1 %>% mutate(localO.h = ifelse(localHoW %in% 129:137, "1", ifelse(localHoW %in% 105:113, "1", ifelse(localHoW %in% 81:89, "1", ifelse(localHoW %in% 57:65, "1", ifelse(localHoW %in% 33:41, "1", "0")))))) %>% mutate(localHoD = ifelse(localHoW %in% 0:23, (localHoW %% 24) + 24, ifelse(localHoW %in% 144:167, local(HoW %% 24) + 24, localHoW %% 24))) 
#Author 2 (local):
author2 <- author2 %>% mutate(localO.h = ifelse(localHoW %in% 129:137, "1", ifelse(localHoW %in% 105:113, "1", ifelse(localHoW %in% 81:89, "1", ifelse(localHoW %in% 57:65, "1", ifelse(localHoW %in% 33:41, "1", "0")))))) %>% mutate(localHoD = ifelse(localHoW %in% 0:23, (localHoW %% 24) + 24, ifelse(localHoW %in% 144:167, local(HoW %% 24) + 24, localHoW %% 24))) 
#Author 3 (local):
author3 <- author3 %>% mutate(localO.h = ifelse(localHoW %in% 129:137, "1", ifelse(localHoW %in% 105:113, "1", ifelse(localHoW %in% 81:89, "1", ifelse(localHoW %in% 57:65, "1", ifelse(localHoW %in% 33:41, "1", "0")))))) %>% mutate(localHoD = ifelse(localHoW %in% 0:23, (localHoW %% 24) + 24, ifelse(localHoW %in% 144:167, local(HoW %% 24) + 24, localHoW %% 24))) 
#Author 4 (local):
author4 <- author4 %>% mutate(localO.h = ifelse(localHoW %in% 129:137, "1", ifelse(localHoW %in% 105:113, "1", ifelse(localHoW %in% 81:89, "1", ifelse(localHoW %in% 57:65, "1", ifelse(localHoW %in% 33:41, "1", "0")))))) %>% mutate(localHoD = ifelse(localHoW %in% 0:23, (localHoW %% 24) + 24, ifelse(localHoW %in% 144:167, local(HoW %% 24) + 24, localHoW %% 24))) 
#Author 5 (local):
author5 <- author5 %>% mutate(localO.h = ifelse(localHoW %in% 129:137, "1", ifelse(localHoW %in% 105:113, "1", ifelse(localHoW %in% 81:89, "1", ifelse(localHoW %in% 57:65, "1", ifelse(localHoW %in% 33:41, "1", "0")))))) %>% mutate(localHoD = ifelse(localHoW %in% 0:23, (localHoW %% 24) + 24, ifelse(localHoW %in% 144:167, local(HoW %% 24) + 24, localHoW %% 24))) 
```

#Apache Hadoop Developer Working Hour Patterns

**Heat Map Description:**

- The y-axis in these visualizations represent the hour of day (HoD) when the commit was made. 0-23 are the aggregate hours of the work week, and 24-47 are the aggregate hours for the weekend (i.e. the cell at (2013, 12) represents the total number of commits made in 2013 at 12:00 on a work day).  The x-axis represents the year the commit was made in.  This design choice was suggested by Chris (and I believe originated from Christian) because we want to examine their overall work pattern (do they work during office hours or non-office hours?), and this allows us to see the difference while keeping the scale manageable (the alternative would of been to have all the hours of the week 0-168 and this seemed much less digestible as a visualization).

**Line Graph Description:**

- For these visualizations, the focus is again on the top 5 contributors, except now the visualization will be looking at the number of commits during office hours (9-5 on weekdays) vs non-office hours (non-9-5/weekends).  I chose to visualize this plot with a line graph, where the x-axis is the year and the y-axis is the number of commits.  Both office and non-office hours are plotted on the same graph but in different colors, to help streamline the visualization.

- First I created a new office hour variable (o.h), which will be used for grouping in line graph, 1 = office hours, 0 = non-office hours.  I considered office hours to be from 09:00 - 17:00.  However it should be noted potential time zone issues have not yet been addressed.  

## Contributor 1: 

```{r, echo=FALSE}
#Funtion takes in an author and creates line line graph plotting office hours vs. non-office hours and longitudinal Heat Map of Working Hours for the author 
create_both_graphs <- function(authori) {
     # Aggregates frequencies and groups by o.h and year
    ai <- authori %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)
    
    # Changes o.h from continuous to discrete column 
    ai$localO.h <- as.factor(ai$localO.h)
    
    # Creates line graph plotting office hours vs. non-office hours
    plot(ggplot(ai, aes(x = year, y = freq)) + geom_line(aes(colour= localO.h)) + labs(title = "Office Hour vs. Non-Office Hour Commits\n", x = "Year", y = "Number of Commits", color = "Time of Commit") +
      scale_color_manual(labels = c("During Non-Office Hours", "During Office Hours"), values = c("blue", "red")))
    
    # Aggregates frequencies and groups by HoD
    aib <- authori %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, localHoD, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)
    
    # Creates longitudinal Heat Map of Working Hours:
    plot(ggplot(aib, aes(year, localHoD)) + geom_tile(aes(fill = freq), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue") + scale_y_continuous(breaks=seq(0, 168, 2)))
}
```
```{r}
create_both_graphs(author1)
create_both_graphs(author2)
```



## Contributor 2:
```{r, echo=FALSE}
# Aggregates frequencies and groups by o.h and year
a2 <- author2 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Changes o.h from continuous to discrete column 
a2$localO.h <- as.factor(a2$localO.h)

# Creates line graph plotting office hours vs. non-office hours
ggplot(a2, aes(x = year, y = freq)) + geom_line(aes(colour= localO.h)) + labs(title = "Office Hour vs. Non-Office Hour Commits\n", x = "Year", y = "Number of Commits", color = "Time of Commit") +
  scale_color_manual(labels = c("During Non-Office Hours", "During Office Hours"), values = c("blue", "red"))

# Aggregates frequencies and groups by HoD
a2b <- author2 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, localHoD, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Creates longitudinal Heat Map of Working Hours:
ggplot(a2b, aes(year, localHoD)) + geom_tile(aes(fill = freq), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue") + scale_y_continuous(breaks=seq(0, 168, 2))
```

## Contributor 3:
```{r, echo=FALSE}
# Aggregates frequencies and groups by o.h and year
a3 <- author3 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Changes o.h from continuous to discrete column 
a3$localO.h <- as.factor(a3$localO.h)

# Creates line graph plotting office hours vs. non-office hours
ggplot(a3, aes(x = year, y = freq)) + geom_line(aes(colour= localO.h)) + labs(title = "Office Hour vs. Non-Office Hour Commits\n", x = "Year", y = "Number of Commits", color = "Time of Commit") +
  scale_color_manual(labels = c("During Non-Office Hours", "During Office Hours"), values = c("blue", "red"))

# Aggregates frequencies and groups by HoD
a3b <- author3 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, localHoD, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Creates longitudinal Heat Map of Working Hours:
ggplot(a3b, aes(year, localHoD)) + geom_tile(aes(fill = freq), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue") + scale_y_continuous(breaks=seq(0, 168, 2))
```

## Contributor 4:
```{r, echo=FALSE}
# Aggregates frequencies and groups by o.h and year
a4 <- author4 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Changes o.h from continuous to discrete column 
a4$localO.h <- as.factor(a4$localO.h)

# Creates line graph plotting office hours vs. non-office hours
ggplot(a4, aes(x = year, y = freq)) + geom_line(aes(colour= localO.h)) + labs(title = "Office Hour vs. Non-Office Hour Commits\n", x = "Year", y = "Number of Commits", color = "Time of Commit") +
  scale_color_manual(labels = c("During Non-Office Hours", "During Office Hours"), values = c("blue", "red"))

# Aggregates frequencies and groups by HoD
a4b <- author4 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, localHoD, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Creates longitudinal Heat Map of Working Hours:
ggplot(a4b, aes(year, localHoD)) + geom_tile(aes(fill = freq), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue") + scale_y_continuous(breaks=seq(0, 168, 2))
```

## Contributor 5:
```{r, echo=FALSE}
# Aggregates frequencies and groups by o.h and year
a5 <- author5 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Changes o.h from continuous to discrete column 
a5$localO.h <- as.factor(a5$localO.h)

# Creates line graph plotting office hours vs. non-office hours
ggplot(a5, aes(x = year, y = freq)) + geom_line(aes(colour= localO.h)) + labs(title = "Office Hour vs. Non-Office Hour Commits\n", x = "Year", y = "Number of Commits", color = "Time of Commit") +
  scale_color_manual(labels = c("During Non-Office Hours", "During Office Hours"), values = c("blue", "red"))

# Aggregates frequencies and groups by HoD
a5b <- author5 %>% mutate(freqYr = frequency(year)) %>% group_by(freqYr, localHoD, year, localO.h) %>% summarise(freq = sum(freqYr)) %>% group_by(localO.h)

# Creates longitudinal Heat Map of Working Hours:
ggplot(a5b, aes(year, localHoD)) + geom_tile(aes(fill = freq), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue") + scale_y_continuous(breaks=seq(0, 168, 2))
```

```{r}

a1c <- author1 %>% mutate(freqO.hCommits = frequency(localO.h)) %>% mutate(freqNonO.hCommits = frequency(localO.h)) 
a1c <- a1c %>% group_by(localO.h, )

View(a1c)

```